\vspace{-2.0cm}The effect of nitrogen application time on infection and development of the Wheat Streak Mosaic Virus (WSMV)

### 2015-09-30 draft1leslieandkenny.Rnw
Changed code to glm. 3rd try.

@@

# Combine nitrogen and inoc/cntl (for the heatmap labels)
fert2$n.status <- with(fert2, interaction(n.trt, status))
@

@@

<<fakepre,echo=FALSE>>=
# Create a row factor
fert.sim1$row <- with(fert.sim1, interaction(variety, block))

# Correctly order nitrogen application levels, the default is alphabetical
fert.sim1$n.trt <- relevel(fert.sim1$n.application, 'FALL')

@@

\begin{figure}
\caption{Example heatmap visualization for identifying patterns of zeros.}
<<heatmap,echo=FALSE,message=FALSE>>=
## Heatmap with meaningful ordering

# Order the dataset, using the arrange() function from the dplyr package
require(dplyr)
fert.ordered <- arrange(fert.sim1, variety, block, status, n.trt)

@@

      # Use black for 0, and use heatmap colors (red-orange-yellow-white) for 1 to 30
      col = c('black', heat.colors(30)), zlim = c(0, 30),
      xlab = '', ylab = '', yaxt = 'n', xaxt = 'n') # Don't automatically create axes or labels

@@

\noindent If information about the physical layout of the field is
available, it would be possible to construct a similar heatmap to look
for spatial patterns. If correlation between nearby plots is seen,
this should be included in the model.

@@

\begin{figure}
\caption{Example \texttt{itableplot} output.}
<<tabplot,echo=FALSE,warning=FALSE,message=FALSE>>=
tableplot(fert.ordered, select = c(infected, variety, block, n.trt, status),
          nBins = 30, sortCol = variety)
@
\end{figure}

\subsection{Model Fitting and Possible Issues}
\noindent The
<<models1,warning=FALSE,results='hide',cache=TRUE>>=
glmm3way.sim <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|block/row),
                      control = glmerControl(optimizer = 'bobyqa', optCtrl = list(maxfun = 20000)),
                      family = binomial, data = fert.sim1)
@

<<models2,warning=FALSE,results='hide',cache=TRUE>>=
noVar1.sim <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|block/row),
                      control = glmerControl(optimizer = 'bobyqa', optCtrl = list(maxfun = 20000)),
                      family = binomial, data = fert.sim1, subset = variety!='Var1')
@

<<models3,warning=FALSE,results='hide',cache=TRUE>>=
glmm3way.sim <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|block/row),
                      control = glmerControl(optimizer = 'bobyqa', optCtrl = list(maxfun = 20000)),
                      family = binomial, data = fert.sim1)
@

<<models4,echo=FALSE,warning=FALSE,results='hide',cache=TRUE>>=
glmm3way <- glmer(cbind(infected, total) ~ variety*n.trt*status2 + (1|block/row),
                  control = glmerControl(optimizer = 'bobyqa'),
                  family = binomial, data = fert2)
#summary(glmm3way)

glmm2way <- glmer(cbind(infected, total) ~ variety*n.trt + variety*status2 +
                    n.trt*status2 + (1|block/row),
                    control = glmerControl(optimizer = 'bobyqa'),
                  family = binomial, data = fert2)
#summary(glmm2way)

#Anova(glmm3way)
#Anova(glmm2way)
#anova(glmm2way, glmm3way)
@

<<plots1,eval=FALSE,echo=FALSE,warning=FALSE,message=FALSE,out.height='0.5\\linewidth',fig.height=5>>=
par(mfrow = c(1, 2))
plot(resid(glmm3way) ~ predict(glmm3way), col = variety, data = fert.sim1)
plot(resid(glmm3way) ~ fitted(glmm3way), col = variety, data = fert.sim1)
legend('topright', pch = 1, col = 1:5,
       legend = levels(fert.sim1$variety))
@

<<nosnms,cache=TRUE>>=
table(fert2$infected==0,fert2$variety)

@@

\begin{figure}
\caption{The fitted model can be nicely summarized with plots from the \texttt{effects} package.}
<<plots2,echo=FALSE,warning=FALSE,message=FALSE,out.height='0.5\\linewidth',fig.height=5>>=
require(effects)
plot(allEffects(glmm3way.sim), multiline = TRUE, type = 'response', se = TRUE,
     ci.style='bars', x.var = 'n.trt', rotx = 45, layout = c(5, 1),
     main = 'Variety:Treatment:Inoculation Effects Plot',
     xlab = 'Nitrogen Application Date', ylab = 'Infection Probability',
     key.args = list(title = 'Inoculation Status'))
@
\end{figure}

@@

\subsection{Data Preprocessing}
\noindent We created variable with a unique level for each row, since row
it was not included in the dataset. We also changed the baseline levels
of \verb|n.application| and \verb|status|.
<<fakepre,eval=FALSE>>=
@

>>>>

# If YSTN is the variety we want to know about the most, make it baseline
fert2$variety2 <- relevel(fert2$variety, 'YSTN')

@@

# Combine nitrogen and inoc/cntl (for the heatmap labels)
#fert2$n.status <- with(fert2, interaction(n.trt, status))
@

@@

<<fakepre,echo=FALSE>>=
# Correctly order nitrogen application levels, the default is alphabetical
fert.sim1$n.trt <- relevel(fert.sim1$n.application, 'FALL')

@@

\noindent Any treatment or variety with many black squares is a reason
for concern. In the example, all but one of the inoculated fall-application
plots for Var1 had counts of 0. The software will have difficulty estimating
the the three-way interaction involving fall nitrogen application and
variety 1.

\noindent If a treatment has many zeros across all varieties or a variety
has many zeros, Nar should consider whether he has reason to believe
that treatment or variety has qualitative differences from the others
that would cause it to have a much lower infrection rate. These data would
be considered as coming from a different population than the data with
fewer zeros. \textbf{If and only if it is justified}, Nar could omit all
of the data for this treatment or variety when fitting the model. The scope
of inference for this model must then be limited to varieties and
treatments that have a nonzero infection rate. A separate analysis would
be needed for more resistant population.

\begin{figure}[b]
\caption{Example heatmap visualization for identifying patterns of zeros.}
<<heatmap,echo=FALSE,message=FALSE>>=
## Heatmap with meaningful ordering

# Order the dataset, using the arrange function from the dplyr package
require(dplyr)
fert.ordered <- arrange(fert.sim1, variety, block, status, n.trt)

@@

      # Use black for 0, and use heatmap colors (red-orange-yellow-white) for 1 to 30
      col = c('black', heat.colors(30)), zlim = c(0, 30),

      # Don't automatically create axes or labels
      xlab = '', ylab = '', yaxt = 'n', xaxt = 'n')

@@

\noindent If information about the physical layout of the field is
available, it would be possible to construct a similar heatmap to look
for spatial patterns. If correlation between nearby plots is seen,
the model should be modified to account for this.

@@

\begin{figure}
\caption{Example \texttt{itableplot} output.}
<<tabplot,echo=FALSE,warning=FALSE,message=FALSE>>=
tableplot(fert.sim1, select = c(infected, variety, block, n.trt, status),
          nBins = 30, sortCol = variety)
@
\end{figure}

\subsection{Model Fitting and Possible Issues}

The following code will fit the three-way interaction model:

<<startingmodel,warning=FALSE,results='hide',cache=TRUE>>=
glm3way <- glm(cbind(infected, total) ~ block/variety + variety*n.trt*status2,
               family = quasibinomial, data = fert.sim1)
@

If it is decided that certain observations come from an extremely
resistant population and can be omited from the current analysis.

<<novar1,warning=FALSE,results='hide',cache=TRUE>>=
noVar1 <- glm(cbind(infected, total) ~ block/variety + variety*n.trt*status2,
              family = quasibinomial, data = fert.sim1, subset = variety!='Var1')
@

If there is no interest in comparing between inoculated plots and
control plots, separate models could be fit.

<<separatemodels,warning=FALSE,results='hide',cache=TRUE>>=
onlyINOC <- glm(cbind(infected, total) ~ block/variety + variety*n.trt,
                family = quasibinomial, data = fert.sim1, subset = status=='INOC')

onlyCNTL <- glm(cbind(infected, total) ~ block/variety + variety*n.trt,
                family = quasibinomial, data = fert.sim1, subset = status=='CNTL')
@

<<models5,echo=FALSE,warning=FALSE,results='hide',cache=TRUE>>=
glm3way2 <- glm(cbind(infected, total) ~ block/variety + variety*n.trt*status2,
                family = binomial, data = fert2)
summary(glm3way2)

glm2way2 <- glm(cbind(infected, total) ~ block/variety + variety*n.trt + variety*status2 + n.trt*status2,
                family = quasibinomial, data = fert2)
summary(glm2way2)

Anova(glm3way2, test.statistic = 'F')
Anova(glm2way2, test.statistic = 'F')
Anova(glm2way2, glm3way2, test.statistic = 'F')
anova(glm2way2, glm3way2)
@

<<nosnms,echo=FALSE,warning=FALSE,results='hide',cache=FALSE>>=
table(fert2$infected==0,fert2$variety)

@@

\begin{figure}
\caption{The fitted model can be nicely summarized with plots from the \texttt{effects} package.}
<<plots2,echo=FALSE,warning=FALSE,message=FALSE,out.height='0.5\\linewidth',fig.height=5>>=
require(effects)
plot(Effect(c('variety','n.trt','status2'), glm3way),
     multiline = TRUE, type = 'response', se = TRUE, ci.style='bars',
     x.var = 'n.trt', rotx = 45, layout = c(5, 1),
     main = 'Variety/Treatment/Inoculation Effects Plot',
     xlab = 'Nitrogen Application Date', ylab = 'Infection Probability',
     key.args = list(title = 'Inoculation Status'))
@
\end{figure}

@@

\subsection{Data Preprocessing}
We changed the baseline levels of \verb|n.application| and \verb|status|.
If the YLST variety is the primary variety of interest, the \verb|relevel|
function could be used on the \verb|variety| variable the same way.
<<fakepre,eval=FALSE>>=
@
